Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2018-06-10 paper/referee_reply1.txt

We thank the referee for a thoughtful, expert critique of our paper. We respond to each major and minor comment below, omitting a direct reply when the changes were straightforward to implement.

@@

Comments to the Author
This paper presents size-dependent clustering of SDSS galaxies, and interprets it using a simple model for the dependence of effective radius on halo radius. It is well presented and the results are interesting and useful. I have however a few suggestions/questions that may help to further improve the clarity and strengthen the analysis.

MAJOR:

1) The size-halo connection that the authors advocate, presented in Section 3.2.2, sets each galaxy to be 1% of the size of its halo at the time of Mpeak. The authors state below Eq 3 that this defines a _median_ relation between Reff and Mhalo, around which they assume a log-normal scatter of 0.2 dex. However, it seems that a more direct application of Eq 2 would be to apply it on a halo-by-halo basis, since Rpeak is known for every object in the catalogue. This would specify the full Reff distribution, with scatter deriving from the scatter in Rpeak. Why is this not done? No justification is given for using a log-normal scatter of 0.2 dex.

Reply) The concluding sentence to Section 3.2.2 was indeed confusing; it turns out it was also misleading: in fact, in our analysis we assigned R_1/2 in exactly the fashion the referee described, about each halo's own R_Mpeak. We have changed the language to more accurately reflect the implementation of the model, and thank the referee for pointing out this confusing language. We have also expanded the discussion for our choice of the fiducial level of scatter, as described further in our reply to the next comment.

### 2018-06-14 paper/referee_reply1.txt

We thank the referee for a thoughtful, expert critique of our paper. We respond to each major and minor comment below, omitting a direct reply to the minor comments when the changes were straightforward to implement.

@@

Reply) We agree with the referee that drawing directly from P_SDSS(R_1/2 | M*) would improve the fidelity of our model with respect to the SDSS. However, our goal here is actually twofold. On the one hand, we want to identify a model that has sufficient flexibility to capture size-dependent clustering with good accuracy, using SDSS as our validation dataset. On the other hand, we are attempting to build a class of models that can be applied equally well to other datasets, including those may not have the statistics of the SDSS, as is the case in present-day high-redshift surveys. We have added what we hope is a clarifying paragraph on this point towards the end of Section 7.

3) Why use Rvir at Mpeak and not at any other time (say z=0)? I assume this is simply because Mpeak gives the best match to the observations, but this is not explained and hence comes across as somewhat arbitrary. If the result is sensitive to this (as suggested in the final paragraph of Section 5.2), then clustering provides a means to distinguish different choices for halo size in terms of how well they predict Reff. Do the authors know which choices would or would not work? It would be helpful to compare the clustering predicted by one or more alternatives.

Reply) Our ``tidal stripping" model is an attempt to explore the consequences of using Mvir(z=0) to model galaxy size. There are certainly many other possible choices one can make for the halo property used to connect to size, such as the mass at the time of first infall, at the time of last infall, etc. We agree with the referee that a more comprehensive study of the hyper-parameter space is warranted by our results. In particular, we have found one good-fitting model based on R_Mpeak, and several other poor-fitting models based on alternative choices for the galaxy-halo connection. This directly implies that observations of galaxy clustering can indeed discriminate between such choices. Thus we consider a comprehensive study (analogous to the Reddick+13 SHAM study) well-motivated by our work, but beyond the present scope.

@@

4) There should be an uncertainty on w_p in the mock data, i.e. on the lines in Figs. 4-6. Since the model is probabilistic (both the abundance matching step and size assignment have random scatters), different Monte Carlo realisations of the model should give different results. Further, for a given realisation jackknife resampling would give an uncertainty estimate in a similar manner to the SDSS data. I would like to see these uncertainties as bands on the plots, or at least a general estimate of their magnitude. Looking at the bottom-right panel of Fig. 6, for the green line to be consistent with 0 the uncertainty on it must be fairly large, at least at low r; without this uncertainty it is hard to gauge the significance of the difference between the purple line and the data.

Reply) We have added jackknife error bars in the form of shaded bands surrounding wp in the figures comparing models to data.

@@

7) Final paragraph of Section 6: It has been known for almost 20 years that the scatter in halo spin at fixed mass is larger than the scatter in galaxy size (de Jong & Lacey 2000, Gnedin et al 2007). The forward modelling implications of Reff^{disk} \propto \lambda have already been explored by Desmond & Wechsler 2015, who find it to be problematic for the Tully-Fisher relation as well as the scatter in size. Further, most modern hydrodynamical simulations (including Illustris [Rodriguez-Gomez et al 2017], EAGLE [Sales et al 2009, Desmond et al 2017] and Magneticum [Teklu et al 2015]) correlate Reff little if at all with halo spin. (It would be useful in general to provide a qualitative comparison with hydro sims, which have the potential to test models for the size-halo connection from first principles, and are beginning to do so.) Again it is unclear to me why Somerville et al 2017 should be of any particular importance for the scatter: The scatter in Reff at fixed M* comes straight from the data, and the scatter in lambda at fixed Mhalo (only relevant for angular momentum-based size models, which are barely mentioned here) comes straight from the N-body sim.

Reply) We have added a new sentence that clarifies the purpose of the final paragraph of Section 6, which is to call attention to the methodological similarity of our model to that in Somerville+17. We are interested in applying our forward-modeling techniques to see what insight can be learned about the R_1/2^{disk} \propto \lambda from galaxy clustering, which would be complementary to the knowledge gained by the Desmond & Wechsler 2015 study.

Studying hydro sims with two-point statistics is the subject of ongoing work. We are including, for the referee's interest, a plot showing the results of a recent calculation of the clustering ratios exhibited by Illustris galaxies. We consider these and related results more appropriate for a separate follow-up paper.

We note that although the correlation of size or baryon specific angular momentum with current specific halo angular momentum is found to be weak in current simulations, some simulations show a nearly linear size-virial radius correlation consistent with observations (J. Schaye, private comm.). Also, the observed distribution of galaxy sizes is comparable to that expected in simple spin-based models from the halo spin distribution, when galaxies of *all* types are considered. Thus, the smaller scatter of sizes exhibited by disk galaxies may indicate either that sizes are affected by other factors and not only by spin, or, alternatively, that disk galaxies do not sample the entire spin distribution of halos. Both possibilities are very interesting and motivate future studies, which we believe are outside the scope of this paper.

### 2018-06-14 paper/galsize_paper.tex
Completed revisions after implementing changes suggested by co-authors

%%%%%%%%  galaxy properties  %%%%%%%%
\newcommand{\rhalf}{R_{1/2}}
\newcommand{\rhalfdisk}{R_{1/2}^{\rm disk}}
\newcommand{\rhalfbulge}{R_{1/2}^{\rm bulge}}
\newcommand{\adisk}{A_{\rm disk}}
\newcommand{\abulge}{A_{\rm bulge}}
\newcommand{\alphadisk}{\alpha_{\rm disk}}
\newcommand{\alphabulge}{\alpha_{\rm bulge}}
\newcommand{\sigmarhalf}{\sigma_{\rm R_{1/2}}}
\newcommand{\bt}{{\rm B/T}}
\newcommand{\mstar}{M_{\ast}}
\newcommand{\ssfr}{{\rm sSFR}}
\newcommand{\sfr}{{\rm SFR}}
\newcommand{\fsat}{F_{\rm sat}}

@@

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=9cm]{FIGS/sdss_size_vs_mstar_scatter.pdf}
\includegraphics[width=\textwidth]{FIGS/sdss_small_large_sample_definitions.pdf}
\caption{
{\bf Definition of ``small" and ``large" galaxies.} For a volume-limited SDSS galaxy sample defined by $M_{\ast}>10^{10.25}M_{\odot},$ the {\em top panel} visually demonstrates how we classify galaxies into ``small" and ``large" subsamples in an $\mstar$-dependent manner. As described in detail in \S\ref{subsec:sizedef}, we compute the median value $\median{\rhalf}{\mstar}$ at the stellar mass of each galaxy in the sample, and divide the sample into two based on this $\mstar-$dependent cut. The {\em bottom left panel} compares the stellar mass functions $\phi(\mstar)$ of the two samples, confirming that our method for separating ``small" from ``large" galaxies yields subsamples with statistically indistinguishable $\phi(\mstar)$. The {\em bottom right panel} shows histograms of $\rhalf$ for the two subsamples, which partially overlap due to the stellar mass range spanned by the sample.
}
\label{fig:sizedefinition}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

@@

It is natural to consider the possibility that stellar mass is stripped from satellite galaxies after infall. The basis for our implementation of this phenomenon is the fitting function presented in \citet{smith_etal16}, which was calibrated by studying stellar mass loss in a suite of high-resolution hydrodynamical simulations. In this model, $f_{\ast}$ quantifies the remaining fraction of stellar mass as a function of $f_{\rm DM},$ the fraction of dark matter mass that remains after subhalo infall:
\beq
f_{\ast} = 1 - \exp(-14.2f_{\rm DM}).
\eeq
For $f_{\rm DM}$ we use the ratio of present-day subhalo mass divided by the peak mass, $M_{\rm vir}/M_{\rm peak}.$ If we denote the post-stripping stellar mass as $M_{\ast}',$ then we have $M_{\ast}'\equiv f_{\ast}M_{\ast},$ where $M_{\ast}$ is given by the initial application of abundance matching. We then calculate the post-stripping radius by interpolating $\langle\rhalf'\vert\mstar'\rangle$ directly from SDSS data. In this way, we model satellite mass loss and  size truncation in a manner that mimics what is seen in hydrodynamical simulations. As shown in \citet{smith_etal16}, stellar stripping in this model does not begin until $f_{\rm DM}\lesssim0.2,$ and in order for $\sim50\%$ of stars to be stripped, the DM halo needs to be stripped to $\sim5\%$ of its initial mass. We note that mass loss slightly depletes the stellar mass function of this model relative to the \citet{moustakas_etal13}, but the differences are less than $10\%$ for stellar masses $\mstar\gtrsim10^{9.5}\msun.$

@@

In the $\rvir-$based model, satellites and splashback galaxies are smaller than centrals of the same halo mass due to the physical size of their halo being smaller at earlier times $\zpeak.$ There are two distinct reasons why this feature results in small galaxies being more strongly clustered relative to larger galaxies of the same mass. First and foremost, satellite galaxies statistically occupy higher mass host halos that are more strongly clustered. So in models where satellites are smaller than centrals of the same mass, smaller galaxies will have a higher satellite fraction, boosting the clustering of small relative to large galaxies of the same stellar mass. Second, at fixed mass, halos of $L_\ast$ galaxies that form earlier are more strongly clustered, a phenomenon commonly known as {\em halo assembly bias} \citep{gao_white05,wechsler_etal06}. Splashback halos are typically earlier-forming than true centrals \citep{wang_etal09}, and so models where splashback halos host smaller-than-average galaxies will naturally predict smaller galaxies being the more strongly clustered \citep[see, e.g.,][for a demonstration of the splashback-dependence of halo clustering]{sunayama_etal16}.

@@

Recent results indicate that some portion of subhalos may be artificially disrupted \citep{guo_white13, campbell_etal17}. To address this possibility in our modeling, we use an extension of {\tt Consistent Trees} that models the evolution of subhalos after disruption. The phase space evolution of disrupted subhalos is approximated by following a point mass evolving in the host halo potential according to the orbital parameters of the subhalo at the time of disruption; the evolution of subhalo mass and circular velocity is approximated using the semi-analytic model presented in \citet{jiang_vdB14}. We then use the {\tt extract\_orphans} program in {\tt UniverseMachine} to walk through all the Bolshoi-Planck {\tt hlist} files, yielding the main progenitor information of every subhalo that was ever identified by {\tt Consistent Trees}.

>>>>

%%%%%%%%  galaxy properties  %%%%%%%%
\newcommand{\rhalf}{R_{1/2}}
\newcommand{\rhalfdisk}{R_{1/2}^{\rm disk}}
\newcommand{\rhalfbulge}{R_{1/2}^{\rm bulge}}
\newcommand{\adisk}{A_{\rm disk}}
\newcommand{\abulge}{A_{\rm bulge}}
\newcommand{\alphadisk}{\alpha_{\rm disk}}
\newcommand{\alphabulge}{\alpha_{\rm bulge}}
\newcommand{\sigmarhalf}{\sigma_{\rm R_{1/2}}}
\newcommand{\bt}{{\rm B/T}}
\newcommand{\mstar}{M_{\star}}
\newcommand{\ssfr}{{\rm sSFR}}
\newcommand{\sfr}{{\rm SFR}}
\newcommand{\fsat}{F_{\rm sat}}

@@

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=9cm]{FIGS/sdss_size_vs_mstar_scatter.pdf}
\includegraphics[width=\textwidth]{FIGS/sdss_small_large_sample_definitions.pdf}
\caption{
{\bf Definition of ``small" and ``large" galaxies.} For a volume-limited SDSS galaxy sample defined by $M_{\star}>10^{10.25}M_{\odot},$ the {\em top panel} visually demonstrates how we classify galaxies into ``small" and ``large" subsamples in an $\mstar$-dependent manner. As described in detail in \S\ref{subsec:sizedef}, we compute the median value $\median{\rhalf}{\mstar}$ at the stellar mass of each galaxy in the sample, and divide the sample into two based on this $\mstar-$dependent cut. The {\em bottom left panel} compares the stellar mass functions $\phi(\mstar)$ of the two samples, confirming that our method for separating ``small" from ``large" galaxies yields subsamples with statistically indistinguishable $\phi(\mstar)$. The {\em bottom right panel} shows histograms of $\rhalf$ for the two subsamples, which partially overlap due to the stellar mass range spanned by the sample.
}
\label{fig:sizedefinition}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

@@

It is natural to consider the possibility that stellar mass is stripped from satellite galaxies after infall. The basis for our implementation of this phenomenon is the fitting function presented in \citet{smith_etal16}, which was calibrated by studying stellar mass loss in a suite of high-resolution hydrodynamical simulations. In this model, $f_{\star}$ quantifies the remaining fraction of stellar mass as a function of $f_{\rm DM},$ the fraction of dark matter mass that remains after subhalo infall:
\beq
f_{\star} = 1 - \exp(-14.2f_{\rm DM}).
\eeq
For $f_{\rm DM}$ we use the ratio of present-day subhalo mass divided by the peak mass, $M_{\rm vir}/M_{\rm peak}.$ If we denote the post-stripping stellar mass as $M_{\star}',$ then we have $M_{\star}'\equiv f_{\star}M_{\star},$ where $M_{\star}$ is given by the initial application of abundance matching. We then calculate the post-stripping radius by interpolating $\langle\rhalf'\vert\mstar'\rangle$ directly from SDSS data. In this way, we model satellite mass loss and  size truncation in a manner that mimics what is seen in hydrodynamical simulations. As shown in \citet{smith_etal16}, stellar stripping in this model does not begin until $f_{\rm DM}\lesssim0.2,$ and in order for $\sim50\%$ of stars to be stripped, the DM halo needs to be stripped to $\sim5\%$ of its initial mass. We note that mass loss slightly depletes the stellar mass function of this model relative to the \citet{moustakas_etal13}, but the differences are less than $10\%$ for stellar masses $\mstar\gtrsim10^{9.5}\msun.$

@@

In the $\rvir-$based model, satellites and splashback galaxies are smaller than centrals of the same halo mass due to the physical size of their halo being smaller at earlier times $\zpeak.$ There are two distinct reasons why this feature results in small galaxies being more strongly clustered relative to larger galaxies of the same mass. First and foremost, satellite galaxies statistically occupy higher mass host halos that are more strongly clustered. So in models where satellites are smaller than centrals of the same mass, smaller galaxies will have a higher satellite fraction, boosting the clustering of small relative to large galaxies of the same stellar mass. Second, at fixed mass, halos of $L_\star$ galaxies that form earlier are more strongly clustered, a phenomenon commonly known as {\em halo assembly bias} \citep{gao_white05,wechsler_etal06}. Splashback halos are typically earlier-forming than true centrals \citep{wang_etal09}, and so models where splashback halos host smaller-than-average galaxies will naturally predict smaller galaxies being the more strongly clustered \citep[see, e.g.,][for a demonstration of the splashback-dependence of halo clustering]{sunayama_etal16}.

@@

Recent results indicate that some portion of subhalos may be artificially disrupted \citep{guo_white13, campbell_etal17, vdBosch_etal17, vdbosch_ogiya18}. To address this possibility in our modeling, we use an extension of {\tt Consistent Trees} that models the evolution of subhalos after disruption. The phase space evolution of disrupted subhalos is approximated by following a point mass evolving in the host halo potential according to the orbital parameters of the subhalo at the time of disruption; the evolution of subhalo mass and circular velocity is approximated using the semi-analytic model presented in \citet{jiang_vdB14}. We then use the {\tt extract\_orphans} program in {\tt UniverseMachine} to walk through all the Bolshoi-Planck {\tt hlist} files, yielding the main progenitor information of every subhalo that was ever identified by {\tt Consistent Trees}.
